---
title: "Answers for Exercise One - Bar Charts"
---
---
```{r echo=FALSE, message = FALSE}
library(ggplot2)
library(binr)
library(dplyr)
library(DT)
library(tools)
library(RColorBrewer)
library(readxl)
library(reshape2)
library(tidyr)
library (scales)
library(latticeExtra)
library(cowplot)
source("3D_helper.R")
source("piebarHelper.R")
library(knitr)
library(kableExtra)
library(DT)
library(GGally)
library(readxl)
```
<h5>The Data</h5>
```{r echo = FALSE}
  ExerciseOne <-
    data.frame(
      Model = c("Their Model 1", "Their Model 2", "Our Model 1", "Our Model 2"),
      Accuracy = c(0.589415, 0.590705, 0.590867, 0.592849)
    )
head(ExerciseOne)
```

<h5>The Answer</h5>

If your rendered plot looks similar to the plot below, you have successfully created a visualisation that does not mislead your audience. 

Note: You did not need to choose the same color scheme as below. As long as you chose a neutral color scheme and the two different types of models are colored the same then you have correctly completed this exercise. 

```{r echo = FALSE}
  ggplot(data = ExerciseOne, aes(x = Model, y = Accuracy, fill = Model)) + geom_bar(stat ="identity") + coord_cartesian(ylim = c(0, 1)) + scale_fill_manual(values = c("lightskyblue3", "lightskyblue3", "yellow4", "yellow4"))
```
